---
title: "Project 2"
author: "Alexandra Ferentinos"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    latex_engine: xelatex
  output:
    pdf_document:
      latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)       # load libraries you always use here
library(dplyr)
library(scales)

```

## Project 2


#  Is there a significant linear relationship between year built and square footage? 

The histogram visualization notes a right density regarding frequency of homes built. There appears to be a positive trend between year built and frequency of homes built. This implies that there is an upward trend in home construction throughout the years from 1900's to the 2000's. This could be due to an increase in housing construction due to population growth of the area.

This visualization relates to the above question as from this positive trend we can then note if the squarefootage has increased over the years as well, perhaps due to larger home popularity throughout the modern residential construction age, or perhaps other factors. 


```{r}
Data<-read.csv("kc_house_data.csv", sep=",", header=TRUE)
set.seed(6021)
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F)
train<-Data[sample.data, ]
test<-Data[-sample.data, ]


# Create a histogram for 'year built' frequency 
ggplot(train, aes(x = yr_built)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs(title = "Histogram of Year Built",
       x = "Year Built",
       y = "Frequency")
```


#Can we predict whether a house is on a waterfront based on other qualities? Square footage, # bedrooms, # bathrooms, sqft_living15, condition, grade. 

The scatter plot data visualization indicates a concentration of residential properties within the range of 0 to 10 bedrooms in terms of frequency. Additionally, it demonstrates that homes with higher square footage tend to be centered within 0 to 10 bedrooms.. From the data, it seems that a the majority of homes are not waterfront homes. 

It is interesting to note that the outliers in terms of the square footage of the living spaces do not seem to be distinctly influenced by the proximity of a home to a waterfront. This observation may imply that the size of the living area of a house may not necessarily correlate with its location relative to a waterfront area, this could be that other predictors are more influential in the response variable square footage. So, from the waterfront quality in consideration to square footage we cannot necessarily predict whether the house is a waterfront home. 
```{r}
# Create the scatterplot
ggplot(train, aes(x = bedrooms, y = sqft_living, color = as.factor(waterfront))) +
  geom_point() +
  labs(title = "Scatterplot of Number of Bedrooms vs Square Footage ",
       x = "Number of Bedrooms",
       y = "Square Footage",
       color = "Waterfront")
```



