---
title: "project 2"
author: "John Le"
date: "2023-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
kc_house_data <- read.csv('/Users/johnle/Downloads/kc_house_data.csv')
head(kc_house_data)
set.seed(6021)
sample.kc_house_data<-sample.int(nrow(kc_house_data), floor(.50*nrow(kc_house_data)), replace = F)
train<-kc_house_data[sample.kc_house_data, ]
test<-kc_house_data[-sample.kc_house_data, ]
1

```

Scatter Plot between year built and square footage:

The scatter plot visualizes the relationship between the year a house was built ("Year Built") and its living space square footage ("Square Footage of Living Space"). Here's what we can interpret from this graph:

Data Spread: The points are spread across the entire range of years, from the early 1900s to the 2000s, showing that the dataset includes a wide range of house ages.

Trend Over Time: There doesn't appear to be a clear trend or pattern indicating a strong relationship between the year built and the size of the living space. The distribution of square footage is fairly consistent across the years, with a large concentration of homes between 1000 and 3000 square feet regardless of the year built.

Modern Homes: Towards the more recent years (post-1975), there is a visible increase in the number of homes with larger square footage, which could suggest a trend in building larger homes in more recent decades.

Outliers: There are a few outliers, particularly homes with exceptionally large square footage, which stand out above the main cluster of data points. These could represent mansions or unusually large properties.

Older Homes Variation: For the early 1900s, the data points are more sparse, which could be due to fewer homes from that era being included in the dataset, or it could reflect less consistency in house sizes during those years.
```{r}
library(ggplot2)

# Assuming your data frame is named 'kc_house_data' and is already loaded in R
ggplot(kc_house_data, aes(x = yr_built, y = sqft_living)) +
  geom_point(alpha = 0.6) + 
  theme_minimal() +
  labs(title = "Scatter Plot of Square Footage vs Year Built",
       x = "Year Built",
       y = "Square Footage of Living Space")

```

Histograms for each variable split by waterfront status:

The visualization is a histogram of the "sqft_living" variable, which represents the living area square footage of homes, split by waterfront status. The histogram allows us to compare the distribution of living area sizes for homes that are either on the waterfront (denoted by 1) or not on the waterfront (denoted by 0).

From the graph, we can observe the following:

Distribution Shape: The distribution of living area sizes for both waterfront and non-waterfront homes appears right-skewed, indicating that there are a larger number of homes with smaller living areas and fewer homes with larger living areas.

Comparison by Waterfront Status:

Homes that are not on the waterfront (red bars) are more numerous across the range of "sqft_living" sizes. This suggests that non-waterfront homes make up a larger proportion of the data.
Waterfront homes (blue bars) are less frequent, which is expected since waterfront properties are typically rarer and potentially more desirable.
Range and Outliers:

The range of "sqft_living" for non-waterfront homes extends from the smallest to the largest sizes, showing a wide variety of home sizes.
Waterfront homes tend to have larger living areas on average, with fewer small-sized homes compared to non-waterfront homes.
Frequency of Specific Sizes:

There is a peak in frequency for non-waterfront homes at the lower end of the "sqft_living" scale, indicating a high concentration of smaller homes.
For waterfront homes, the data are more evenly distributed, although with a much lower frequency overall due to fewer waterfront homes in the dataset.
```{r}
library(ggplot2)

# Variable of interest
variable <- "sqft_living"

# Create histogram for the variable
ggplot(kc_house_data, aes_string(x = variable, fill = "factor(waterfront)")) +
  geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Waterfront", x = variable, y = "Frequency") +
  theme_minimal() +
  labs(title = paste("Histogram of", variable, "by Waterfront Status"))

```

