---
title: "Project_2_Prelim"
author: "Abner Casillas-colon"
date: "2023-11-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

```{r}
Data <- read.csv("kc_house_data.csv", sep=",", header=TRUE)
Data$waterfront <- factor(Data$waterfront)
set.seed(6021)
sample.data <- sample.int(nrow(Data), floor(.50*nrow(Data)),replace = F)
train <- Data[sample.data,]
test <- Data[sample.data,]
```


```{r}
sqrt(nrow(train))
```
1. Histogram square footage frequency (few observations on one end or the other may skew results) 

The histogram below displays a distribution for the squarefoot living space of the houses within the training data set. Our distribution is right skewed indicating potentially to be on the lookout for outliers on the extreme end that will impact the linearity of the relationship when we conduct our testing. Additionally, if we are to describe our data set we would expect the mean value to be higher than the median and may want to use the median value as a descriptor when discussing how to evaluate the central point of housing living space as a better representation of the data set.
```{r}
ggplot(train, aes(x=sqft_living))+
  geom_histogram(bins=104, fill="cornflowerblue")+
  labs(x="Living Space in sqft", y="Count", title= "Histogram of Housing sqft Living Space")
```

2. Boxplot – density of data at different values (shows distribution of data) - condition or grade based on waterfront or not 

The x axis for this visualization represents grade which for context indicates that 1-3 can be considered "poor construction" 7 has an average quality and 11-13 have a high quality. In this case the "1" group indicates that the property is waterfront while "0" indicates that it is not a waterfront property. When examining the properties without comparing their categories we note that waterfront properties tend to be fairly soundly made averaging a score of approximately 8 and nothing within the poor construction range. The distributions appears to be fairly even around the average score of 8 as well. This is contrasted by non-waterfront properties which have 7 outliers and houses that seem to be lower grade construction averaging around 7. When we conduct the analysis we may wish to keep an eye out for the sheer volume of outliers in the non waterfront property category that may potentially skew the analysis.
```{r}
ggplot(train, aes(y=waterfront, x=grade))+
  geom_boxplot(fill="aquamarine2")+
  labs(x="Grade of the Property", y="Waterfront Status", title="Boxplots of Grade based on if property is Waterfront")
```




