---
title: "Project_2_Data_Visualizations_S4_S6"
author: "Abner Casillas-colon"
date: "2023-11-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```


```{r}
Data <- read.csv("kc_house_data.csv", sep=",", header=TRUE)
Data$waterfront <- factor(Data$waterfront)
Data$view <- factor(Data$view)
set.seed(6021)
sample.data <- sample.int(nrow(Data), floor(.50*nrow(Data)),replace = F)
train <- Data[sample.data,]
test <- Data[sample.data,]
```


S4- Grade (box or density plot)

This barchart of grade of the property reveals a fairly normal distribution with a slight skew to the right. As grade tends to be a representation of housing quality this is something that should be expected as some higher value properties may potentially skew the results to potentially not resemble a normal distribution. For this variable, there appears to be few potential outliers though it may still be worthwhile to examine both sets of outliers when considering if there are any influential obeservations that will impact the model.
```{r}
ggplot(train, aes(x=grade))+
  geom_bar(fill="aquamarine2")+
  labs(x="Grade of the Property",  title="Bar Chart of Grade")
```
s6 - Scatterplot Displaying bedrooms vs bathrooms
```{r}
ggplot(train, aes(x=bedrooms, y=bathrooms))+
  geom_point(fill="aquamarine2")+
  labs(x="Number of Bedrooms", y="Number of Bathrooms", title="Scatterplot of Bedrooms vs Bathrooms")
```
The scatter plot below plots the relationship of number of bedrooms vs bathrooms in the set of sampled homes while also displaying the density of the points to have an idea of where the majority of points lie. We note that the relationship tends to be linear with the majority of observations containing 2-4 bedrooms and 1-3 bathrooms. There maybe a potential influential observation however that has over 30 bedrooms with less than 2 bathrooms. We may wish to see if this observation has other strange properties that will potentially skew the results when conducting further analysis.
```{r}
ggplot(train, aes(x=bedrooms, y=bathrooms))+
  geom_count(fill="aquamarine2")+
  labs(x="Number of Bedrooms", y="Number of Bathrooms", title="Scatterplot of Bedrooms vs Bathrooms")
```





