---
title: "Project 2"
author: "Alexandra Ferentinos"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    latex_engine: xelatex
  output:
    pdf_document:
      latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)       # load libraries you always use here
library(dplyr)
library(scales)

```

## Project 2


#  sale price (response) predictors: sqft, #bedrooms, condition, #floors, sqft above

The box plot visualization of sale price against house condition, ranging from 1 to 5, (5 being the best) shows interesting information about the relationship between condition and sale price. Houses in condition 3 to 5 have the most outliers. This could indicate that there is a higher variance in sale prices within house conditions 3 to 5. 

This visualization relates to the above question as it shows that house condition 3 to 5 has higher variability in sale price meaning there may be other predictors affecting this nuance. So more analysis of sale price(response), to the above predictors from the question could illuminate such outliers. 


```{r}
Data<-read.csv("kc_house_data.csv", sep=",", header=TRUE)
set.seed(6021)
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F)
train<-Data[sample.data, ]
test<-Data[-sample.data, ]

##to assist with the visualization 
log_price<-log10(train$price)
#boxplot of House condition vs Sale Price
ggplot(train, aes(x = factor(condition), y = log_price)) +
  geom_boxplot(fill="green") +
  labs(title = 'Boxplot of House Condition vs Sale Price',
       x = 'House Condition',
       y = 'Sale Price')
```


#Can we predict whether a house is on a waterfront based on other qualities? Square footage, # bedrooms, # bathrooms, sqft_living15, condition, grade. 

The scatter plot data visualization indicates a concentration of residential properties within the range of 0 to 10 bedrooms in terms of frequency. Additionally, it demonstrates that homes with higher square footage tend to be centered within 0 to 10 bedrooms.. From the data, it seems that a the majority of homes are not waterfront homes. 

It is interesting to note that the outliers in terms of the square footage of the living spaces do not seem to be distinctly influenced by the proximity of a home to a waterfront. This observation may imply that the size of the living area of a house may not necessarily correlate with its location relative to a waterfront area, this could be that other predictors are more influential in the response variable square footage. So, from the waterfront quality in consideration to square footage we cannot necessarily predict whether the house is a waterfront home. 
```{r}
# Create the scatterplot
ggplot(train, aes(x = bedrooms, y = sqft_living, color = as.factor(waterfront))) +
  geom_point() +
  labs(title = "Scatterplot of Number of Bedrooms vs Square Footage ",
       x = "Number of Bedrooms",
       y = "Square Footage",
       color = "Waterfront")
```

# Indicators of Mulitcollinearity:
From the correlation matrix values will range from -1, 0, and 1. Where closeness to -1 indicates a strong negative correlation and closeness to 1 indicates a strong positive correlation. As such, closeness to 0 indicates that there is no correlation. 

From the predictors mentioned in Question 1 the following appear to have positive correlations:

-sqft_living and grad (0.756)
-sqft_living and sqft_above (0.874)
-sqft_living and sqft_living15 (0.754)
-bathrooms and sqft_living (0.755)

The following appear to have negative correlations:

-condition and yr_built (-0.366)
-sqft_above and sqft_basement (-0.036)
-condition and sqft_above(-0.159)


```{r}
# Create the scatterplot matrix
library(GGally)
##filter predictors and keep response 
#predictor interaction, with response 
round(cor(train[,c(-1,-2,-8,-9,-10,-17,-18,-19)]),3)

```

#VIFS:
Removed predictor variable sqft_basement as it exhibited low correlatoin to reponse variable price, (.338) This decision was based on the relevancy of the model to predictors and due to the low correlation to price this predictor was dropped. 

From the VIFS it seems that sqft_living has a VIF of 7.305 indicating high levels of multicollinearity. The other VIFs values are below 5, deeming them to be at acceptable levels for multicollinearity threshold. Further investigation of sqft_living will need to be done as it seems to be exhibit multicollinearity characteristics. 

The reduced VIF are now all below 4 suggesting a non issue with multicollinearity. 
```{r}
##VIFS
#view(train)
##dropped sqft_basement
library(faraway)
result_full <- lm(price ~ bedrooms + bathrooms + sqft_living + sqft_lot + condition + grade + sqft_above + yr_built + yr_renovated + sqft_living15 + sqft_lot15, data = train)
vif(result_full)

##reduced model (removing sqft_living)
result_reduced<- lm(price ~ bedrooms + bathrooms + sqft_lot + condition + grade + sqft_above + yr_built + yr_renovated + sqft_living15 + sqft_lot15, data = train)
vif(result_reduced)
```

#Condition:

  The boxplot reveals a discernible trend wherein waterfront properties exhibit a higher prevalence of homes with condition ratings of 4 or 5. Rationale for this phenomenon may lie in the elevated maintenance standards necessitated by the unique environmental challenges associated with waterfront living.

  It is plausible that the proximity of these properties to bodies of water exposes them to heightened levels of humidity, salt, and other elemental factors. In response to these environmental stressors, property owners and builders may find it imperative to implement more stringent maintenance protocols. Such measures would be designed to mitigate the potential damage to the property's siding or foundation, ultimately preserving the overall structural integrity of the house.

  In essence, the observed concentration of houses with condition ratings of 4 or 5 among waterfront properties could be attributed to the conscientious efforts taken to fortify these structures against the deleterious effects of their aquatic/waterfront surroundings. 
```{r}
library(ggplot2)

# Convert to factor, arrange condition
train$condition <- factor(train$condition, levels = c(5,4,3,2,1))
train$waterfront <- factor(train$waterfront)


#colors and arrange from the vector levels
ggplot(train, aes(x = waterfront, fill = condition, group = condition)) +
  geom_bar(position = "fill") +
  labs(title = "Proportion of Condition by Waterfront Status")

```


