---
title: "Project 2 Proposal"
author: "Becky Desrosiers, Abner Castillos, Alexandra Ferentinos, John Le"
date: "2023-11-07"
output: html_document
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = F)
```



# Section 1: Questions of Interest

### Question 1: Is there a significant linear relationship between year built and square footage?

### Question 2: Can we predict whether a house is on a waterfront based on other qualities? Square footage, # bedrooms, # bathrooms, sqft_living15, condition, grade.


The response variable for the first question will be square footage. We are interested in seeing if square footage has increased over time. The response variable for the second data set is waterfront status: is it a waterfront property or not? We are interested in predicting whether a property is waterfront or not based on other properties.


We are interested in the first question because we have noticed that houses are very expensive nowadays. Even so, when we go to historic places and see the types of houses regular people used to live in, it's like what we would call a shack in modern times. Houses have become much more extensive and also include more amenities, such as central heat, air conditioning, and indoor plumbing. We are interested in looking at if square footage can be attributed as one of the factors in the higher prices we see today, by seeing how it has changed over time. Our hypothesis is that it will have increased to a statistically significant degree and is a factor in high housing prices.


We are interested in the second question to find out if there is a significant difference between waterfront properties and non-waterfront properties. We hypothesize that waterfront properties will be higher quality in general (higher square footage, more bedrooms and bathrooms, higher grade), and have neighbors with bigger houses, but perhaps be of lower condition because of the proximity to water and the risk of flooding.

 

# Section 2: Data visualizations 

```{r}
# Split data
Data<-read.csv("kc_house_data.csv", sep=",", header=TRUE)
set.seed(6021)
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F)
train<-Data[sample.data, ]
test<-Data[-sample.data, ]

# Other setup
library(ggplot2)
```


### Question 1: Is there a significant linear relationship between year built and square footage?


```{r}
ggplot(train, aes(x = yr_built)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs(title = "Histogram of Year Built",
       x = "Year Built",
       y = "Frequency")
```

The histogram visualization notes a right skew in the data. There appears to be a positive trend between year built and homes built. This implies that there is an upward trend in home construction throughout the years represented in the dataset. This could be due to an increase in housing construction due to population growth of the area.

This visualization relates to the above question as from this positive trend we can then note if the square footage has increased over the years as well, perhaps due to larger home popularity throughout the modern residential construction age, or perhaps other factors.




```{r}
ggplot(train, aes(x=sqft_living))+
  geom_histogram(bins=104, fill = "skyblue", color = "black")+
  labs(x="Living Space in sqft", y="Count", title= "Histogram of sqft Living Space")
```


The histogram  displays a distribution for the squarefoot living space of the houses within the training data set. Our distribution is right-skewed, indicating potentially to be on the lookout for outliers on the extreme end that will impact the linearity of the relationship when we conduct our testing. Additionally, if we are to describe our data set we would expect the mean value to be higher than the median and may want to use the median value as a descriptor when discussing how to evaluate the central point of housing living space as a better representation of the data set.



```{r}
ggplot(train, aes(x = yr_built, y = sqft_living)) +
  geom_point(alpha = 0.2) + 
  theme_minimal() +
  labs(title = "Scatter Plot of Square Footage against Year Built",
       x = "Year Built",
       y = "Square Footage of Living Space")

```




### Question 2: Can we predict whether a house is on a waterfront based on other qualities? Square footage, # bedrooms, # bathrooms, sqft_living15, condition, grade.



```{r}
ggplot(train, aes(x = bedrooms, y = sqft_living, color = as.factor(waterfront))) +
  geom_point() +
  labs(title = "Scatterplot of Number of Bedrooms vs Square Footage ",
       x = "Number of Bedrooms",
       y = "Square Footage",
       color = "Waterfront")
```


The scatter plot data visualization indicates a concentration of residential properties within the range of 0 to 10 bedrooms in terms of frequency. Additionally, it demonstrates that homes with higher square footage tend to be centered within 0 to 10 bedrooms, with 5 being the most common. From the plot, it appears that a the majority of homes are not waterfront homes. 

It is interesting to note that the outliers in terms of the square footage of the living spaces do not seem to be distinctly influenced by the proximity of a home to a waterfront. This observation may imply that the size of the living area of a house may not necessarily correlate with its location relative to a waterfront area, this could be that other predictors are more influential in the response variable square footage. So, from the waterfront quality in consideration to square footage we cannot necessarily predict whether the house is a waterfront home. It does appear, however, that within a given number of bedrooms, waterfront properties tend to be on the higher end in terms of square footage.

One other important thing to note from this visualization is the outlier with more than 30 bedrooms. This seems very unusual, and may be an error in input, espeically because the square footage is relatively low. We may have to consider dropping this data point.



```{r}
ggplot(train, aes(y= as.factor(waterfront), x = grade))+
  geom_boxplot(fill="skyblue")+
  labs(x="Grade of the Property", y="Waterfront Status", title="Boxplots of Grade for Waterfront and Non-Waterfront Properties")
```




The x-axis for this visualization represents grade which for context indicates that 1-3 can be considered "poor construction" 7 has an average quality and 11-13 have a high quality. On the y-axis, the "1" group indicates that the property is waterfront while "0" indicates that it is not a waterfront property.

When examining the properties without comparing their categories we note that waterfront properties tend to be fairly soundly made averaging a score of approximately 8 and nothing within the poor construction range. The distributions appears to be fairly even around the average score of 8 as well. This is contrasted by non-waterfront properties which have 7 outliers and houses that seem to be lower grade construction averaging around 7. When we conduct the analysis we may wish to keep an eye out for the many outliers in the non-waterfront property category that may potentially skew the analysis.



**Histograms for each variable split by waterfront status**
```{r}
library(ggplot2)

# Variables of interest
variables <- c("sqft_living", "bedrooms", "bathrooms", "condition", "grade")

# Adjust the grid layout based on the number of variables
par(mfrow = c(3, 2))

# Create histograms for each variable
for (var in variables) {
  ggplot(train, aes(x = .data[[var]], fill = factor(waterfront))) +
    geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
    scale_fill_brewer(palette = "Set1") +
    labs(fill = "Waterfront", x = var, y = "Frequency") +
    theme_minimal() +
    labs(title = paste("Histogram of", var, "by Waterfront Status"))
}


```








